---
title: "chapter-8-9-exercise"
author: "Julien Brun"
date: "2/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Yelp Review Data

Exercise:
- Download the Yelp business data as json: https://www.kaggle.com/yelp-dataset/yelp-dataset?select=yelp_academic_dataset_review.json 
- Load the data into Spark
- Check if the data fits in memory on your Standalone instance
- If not, modify the connection configuration to try to make it fit
- Filter the Data for the Restaurants in city of Santa Barbara
- join with the review data
- Do some Exploratory Data Analysis to find your favorite restaurant for next time you come visit :)


Let's start by downloading the business data: `yelp_academic_dataset_business.zip`

## Create the connection
```{r}
library(tidyverse)
library(sparklyr)


# Connect

```

## Load the data
```{r}
# Load the data

```

## Filter Santa Barbara businesses

```{r}

```

## Filter Restaurant businesses

## Create an IDs list to subset the review


## Load the review json

## Subset it