---
title: "R for Data Science Book Club"
subtitle: "Session 3"
author: "John Peach"
institute: "OCRUG"
date: "2012-01-22"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false

---

# Chapter 9 
## Tidy Data with tidyr

1. Every variable must have its own column
1. Each observation must have its own row
1. Each value must have its own cell

That is:
1. Put each dataset in a tibble
1. Put each variable in a column

---
# Chapter 9

.pull-left[
## Gathering

Use when the column names are variables.

**Gather** the columns into rows

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
table4a %>% knitr::kable(format = 'html')
```
]

.pull-right[

```{r, eval=FALSE, echo=TRUE}
table4a %>% 
gather(`1999`, `2000`, 
       key = 'year', 
       value = 'cases')

table4a %>% 
pivot_longer(c(`1999`, `2000`), 
       names_to = 'year', 
       values_to = 'cases')
```

```{r, echo=FALSE}
table4a %>% 
  tidyr::gather(`1999`, `2000`, key = 'year', value = 'cases') %>% 
  knitr::kable(format = 'html')
```
]

---

# Chapter 9

## Spreading

Use when an observation is scattered across muliple rows

**Spread**ing the rows into columns

```{r, echo=FALSE, warning=FALSE, message=FALSE}
table2 %>% knitr::kable(format = 'html')
```


---
# Chapter 9
## Spreading

```{r, eval=FALSE, echo=TRUE}
table2 %>% 
spread(key = type, value = count)

table2 %>% 
pivot_wider(names_from = type, values_from = count)
```

```{r, echo=FALSE}
table2 %>% 
  tidyr::spread(key = type, value = count) %>% 
  knitr::kable(format = 'html')
```

---

# Chapter 9
## Separate
Pulls apart one column into multiple columns

```{r, echo=FALSE}
table3 %>% knitr::kable(format = 'html')
```
---

# Chapter 9
## Separate
```{r, eval=FALSE, echo=TRUE}
table3 %>% 
separate(rate, into=c('cases', 'population'))
```

```{r, echo=FALSE}
table3 %>% 
  tidyr::separate(rate, into=c('cases', 'population')) %>% 
  knitr::kable(format = 'html')
```

---
# Chapter 10
## Relational Data with dpylr

**dplyr** does most of the operations of SQL

1. Inner join
    1. `dplyr::inner_join(<df>, by=<key>)`
1. Outer join
    1. `dpylr::full_join(<df_l>, <df_r>, by=<key>)`
    1. `dpylr::left_join(<df_l>, <df_r>, by=<key>)`
    1. `dpylr::right_join(<df_l>, <df_r>, by=<key>)`
    
`<key>` can be specified as
1. common column name: `by='id'`
1. named vector: `by=c('id' = 'item_id')`
1. natural join: `by=NULL`

---
# Chapter 10
## Filtering Joins

1. `semi_join(x, y)`: keeps all observations in `x` that have a match in `y`
1. `anti_join(x, y)`: drops all observations in `x` that have a match in `y`

* Only columns from `x` are returned
* There is no duplication of rows if `y` has multiple matches

---
# Chapter 11
## Strings with stringr

* `str_c('x', 'y')`: combine string
* `str_sub(x, <start>, <end>)`: subsection
* `str_sort(x, locale = 'en')`: sort
* `str_extract(x, <RegEx>)`: extact
* `str_split(x, <pattern>)`: split
* `str_match(x, <RegEx>)`: returns boolean

---
# Chapter 11
## RegEx

### Metacharacters
* `^`  True at beginning of string or after any \n if /m modifier is used
* `.`  Match a char except \n.  Matches \n if the /m or /i modifiers are used
* `$`  True at end of string or before any \n if /m modifier is used

### Quanifiers
* `*`  Match 0 or more times
* `+`  Match 1 or more times
* `?`  Match 1 or 0 times
* `{COUNT}`  Match exactly COUNT times
* `{MIN,}`   Match at least MIN times
* `{MIN,MAX}` Match at least MIN but not more then MAX times

---
# Questions

* 12.2.1 Q2 - Compute the `rate` for `table2`, and `table4a` ...
* 12.2.2 Q4 - Tidy the simple tibble below. ...
* 13.4.6 Q2 - Add the location of `origin` and `destination` ... 
    * use `library(nycflights13)` tables `flights` and `airports`
* 13.5.1 Q5 - What does `anti_join(flights, airports, by` ...
* 14.2.5 Q3 - Use `str_length()` and `str_sub()` to extract ...
* 14.3.2.1 Q2 - Given the corpus of common words in ...
    * use `library(stringr)` `stringr::words`
* 14.3.4.1 Q3 - Create regular expressions to find all words ...



