---
title: "Advanced R"
subtitle: "Chapters 11-13"
author: "Ted Kyi"
institute: "OCRUG"
date: "2020-10-28"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, intro.css]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      
---
<style>
pre {
    display: block;
    font-family: monospace;
    white-space: pre;
    margin: 1em 0px;
    margin-top: 0em;
    margin-right: 0px;
    margin-bottom: 0em;
    margin-left: 0px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(lobstr)
library(tibble)
library(rlang)
library(purrr)
library(memoise)
library(sloop)
```

# Function operators

* Function operators take functions as input and return a function as output
* A common kind is a wrapper that does a little something extra, then calls the original function
* Examples from `purrr` are `safely()`, `possibly()`, `quietly()`, and `auto_browser()`
* Another example is `memoise::memoise()`
* A common pattern is to call the function operator from a functional, like this:
```{r function operator, echo=TRUE, eval=FALSE}
x <- list(
  c(0.512, 0.165, 0.717),
  c(0.064, 0.781, 0.427),
  c(0.890, 0.785, 0.495),
  "oops"
)
out <- transpose(map(x, purrr::safely(sum)))
```

  
---
# Base types

* Base objects vs. Object-oriented objects
  * OO objects have a `class` attribute
  * `is.object()` returns `TRUE`/`FALSE`
  * `sloop::otype()` returns the object system, like "S3", or "base"*

* Base types aren't an OOP system because primarily use hard coded switch statements in C

* There are 25 (wow!) base types, and objects can have more than one
  * Vectors include NULL, logical, integer, double, etc.
  * Others are function closere, environment
  * S4
  * Others
  

---
# S3

* Useful functions in the `sloop` package:
  * `ftype()` returns the function type, e.g. "S3" "generic"
  * `s3_dispatch()` shows the method called by the S3 generic function
  * Use `s3_get_method()` to see the source code of a S3 method
  * `s3_methods_generic()` lists all methods defined for a generic
  * `s3_methods_class()` lists all methods defined for a class
  
* Recommendation that each S3 class have:  a low-level constructor, a validator, and a user-friendly helper

---
# S3

* In the Object Styles section, showed how classes override `length()`, 
such as data frames are a list of columns, but they return the number of rows instead.

* Inheritance
  * A subclass can delegate work using `NextMethod()`
  * In S3, to allow subclassing, the code for the class needs to be written very carefully.
  For example, if you create a subclass of a tibble, then `x[1]` might return a tibble, not your subclass
  
* Advanced dispatching gets weird, especially for base types and internal generics
  * Infix binary operators do a double dispatch on both arguments,
  then compare results.


---
# Questions

