---
title: "Chap_4-5_exercise_template"
author: "Your Name"
date: "2/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Step 1. Preparation

## 1.1 Load libraries
```{r}
library(sparklyr)
library(ggplot2)
library(dbplot)
library(dplyr)
```

## 1.2 Make a connection to the local spark cluster.
```{r}


```

## 1.3 Load data to the spark cluster. file name: house_price.csv
```{r}


```

### Step 2. EDA

## 2.1 Take a look at the data.
```{r}


```

## 2.2 Check missing value for all features.
```{r}
 


```

## 2.3 Drop the first column of the data.
```{r}


```

## 2.4 Split data into train and test (size = 0.05) set.
```{r}


```

## 2.5 Get numerical summaries of age and distance of the training data.
```{r}


```

### Step 3. Supervised Learning

## 3.1 Fit training data with decision tree.
```{r}


```

## 3.2 Make prediction on test data.
```{r}


```

## 3.3 Evaluate the prediction (report r2).
```{r}


```

### Step 4. Pipelines

## 4.1 Build a transformer for turning all features into vectors.
```{r}


```

## 4.2 Build an estimator that estimates feature mean and standard deviation.
```{r}



```

## 4.3 Create a pipeline with transformer, estimator, and decision tree.
```{r}

```

## 4.4 Fit the pipeline model with training data.
```{r}


```

## 4.5 Make prediction on test data using the pipeline model and evaluate the prediction (report r2).
```{r}

```

## 4.6 Save pipeline model.
```{r}
 


```

## 4.7 Reload pipeline model.
```{r}




```

## 4.8 Extract one estimator from the pipeline.
```{r}



```

## 4.9 Disconnect to the spark cluster
```{r}



```

### Step 5. Unsupervised Learning

## 5.1 Make a connection to the local spark cluster.
```{r}



```

## 5.2 Load `iris` dataset into spark.
```{r}



```

## 5.3 Take a look at the data.
```{r}



```

## 5.4 Check missing value for chosen features (Petal_Length, Petal_Width).
```{r}




```

## 5.4 Check missing value for chosen features (Petal_Length, Petal_Width).
```{r}



```

## 5.5 Fit a model using k-means model with 3 clusters.
```{r}



```

## 5.6 Predict the associated class.
```{r}



```

## 5.7 Plot cluster membership.
```{r}


```

## 5.8 Disconnect to the spark cluster.
```{r}


```

