---
title: "Mastering Spark with R"
subtitle: 'Chapters 4-5'
author: "Lily Cheng"
institute: "Orange County R Users Group"
date: "2021-02-08"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, intro.css]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      
---
<style>
pre {
    display: block;
    font-family: monospace;
    white-space: pre;
    margin: 1em 0px;
    margin-top: 0em;
    margin-right: 0px;
    margin-bottom: 0em;
    margin-left: 0px;
}
</style>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(comment = "")
```

# Intro, Set up, Analysis

* **Spark**: A unified analytics engine for large-scale data processing.
* **Set up**:  `sc <- spark_connect(master = "local", version = "2.3")`
* **Analysis**: `sparklyr` integrates with many R like `dplyr`, `magrittr`, `broom`, `DBI`, `tibble`, `rlang`, and many others

<br>

* However, you cannot use the standard R models. Spark has a large library of modeling commands `MLlib`. 

---

# Exploratory Data Analysis (EDA)



.large[
* **Check for data quality**  
e.g., ` mutate(sex = ifelse(is.na(sex), "missing", sex))` , `glimpse()`

<br>

* **Understand univariate relationships between variables**  
e.g., `sdf_describe()`, `dbplot_histogram()`, `sdf_crosstab()`, `FactoMineR::CA()`

<br>

* **Perform an initial assessment on what variables to include and what transformations need to be done on them**  
]

---

# Feature Engineering

The feature engineering exercise comprises transforming the data to increase the performance of the model.

.large[
* **centering and scaling numerical values**  
e.g., `scaled_age = (age - mean(age)) / sd(age)`

<br>

* **performing string manipulation to extract meaningful variables**  
e.g., convert ethnicity to dummy variables

<br>

* **variable selection**  
e.g., selecting which predictors are used in the model

In addition, save dataset as a Parquet file—an efficient file format ideal for numeric data:

`spark_write_parquet(okc_train, "data/okc-train.parquet")`

]
---

# Supervised Learning
Learning a function that maps an input to an output based on example input-output pairs. It infers a function from labeled training data consisting of a set of training examples.

## Data Split
```{r, eval = FALSE, echo=TRUE, message=FALSE}
data_splits <- sdf_random_split(okc, training = 0.8, testing = 0.2, seed = 42)
okc_train <- data_splits$training
okc_test <- data_splits$testing
```

## Cross-Validation Folds
```{r, eval = FALSE, echo=TRUE, message=FALSE}
vfolds <- sdf_random_split(
  okc_train,
  weights = purrr::set_names(rep(0.1, 10), paste0("fold", 1:10)),
  seed = 42
)
```

---

# Supervised Learning

## Classification
```{r, eval = FALSE, echo=TRUE, message=FALSE}
lr <- ml_logistic_regression(
  analysis_set, not_working ~ scaled_age + sex + drinks + drugs + essay_length
)
```

## Evaluation
```{r, eval = FALSE, echo=TRUE, message=FALSE}
validation_summary <- ml_evaluate(lr, assessment_set)
```

Note: Spark provides evaluation methods for only generalized linear models (including linear models and logistic regression). For other algorithms, you can use the evaluator functions (e.g., ml_binary_classification_evaluator() on the prediction DataFrame) or compute your own metrics. 
Other models, see section 14.5.

---

# Unsupervised Learning
A type of algorithm that learns patterns from untagged data. The hope is that through mimicry, the machine is forced to build a compact internal representation of its world.

## Topic Modeling
```{r, eval = FALSE, echo=TRUE, message=FALSE}
stop_words <- ml_default_stop_words(sc) %>%
  c(
    "like", "love", "good", "music", "friends", "people", "life",
    "time", "things", "food", "really", "also", "movies"
  )
# Latent Dirichlet Allocation
lda_model <-  ml_lda(essays, ~ words, k = 6, max_iter = 1, min_token_length = 4, 
                     stop_words = stop_words, min_df = 5)
```

## Other Models (see section 14.5)
e.g.,
```{r, eval = FALSE, echo=TRUE, message=FALSE}
ml_kmeans() # K-Means Clustering
ml_als() # Alternating Least Squares Factorization
```

---

# Pipelines

Pipelines could be used to express workflow of training a model. We build a pipeline object by providing it with a list of steps.
In spark, the building blocks of pipelines are:

* **Estimator**:  can be used to create a transformer giving some training data
```{r, eval = FALSE, echo=TRUE, message=FALSE}
scaler <- ft_standard_scaler(
  sc,
  input_col = "features",
  output_col = "features_scaled",
  with_mean = TRUE)
```

* **Transformers**: apply transformations to a DataFrame and return another DataFrame; the resulting DataFrame often comprises the original DataFrame with new columns appended to it
```{r, eval = FALSE, echo=TRUE, message=FALSE}
df <- copy_to(sc, data.frame(value = rnorm(100000))) %>% 
  ft_vector_assembler(input_cols = "value", output_col = "features")
```

```{r, eval = FALSE, echo=TRUE, message=FALSE}
scaler_model <- ml_fit(scaler, df)
```

---

# How to create pipelines?

`ml_pipeline()`

## initialize an empty pipeline with ml_pipeline(sc) and append stages to it:
```{r, eval = FALSE, echo=TRUE, message=FALSE}
ml_pipeline(sc) %>% 
  ft_standard_scaler(
    input_col = "features",
    output_col = "features_scaled", 
    with_mean = TRUE)
```

## pass stages directly to ml_pipeline():
```{r, eval = FALSE, echo=TRUE, message=FALSE}
pipeline <- ml_pipeline(scaler)

pipeline_model <- ml_fit(pipeline, df)
pipeline_model
```

---

# Use Case

```{r, eval = FALSE, echo=TRUE, message=FALSE}
cv <- ml_cross_validator(
  sc,
  estimator = pipeline,
  estimator_param_maps = list(
    standard_scaler = list(with_mean = c(TRUE, FALSE)),
    logistic_regression = list(
      elastic_net_param = c(0.25, 0.75),
      reg_param = c(1e-2, 1e-3)
    )
  ),
  evaluator = ml_binary_classification_evaluator(sc, label_col = "not_working"),
  num_folds = 10)
```

```{r, eval = FALSE, echo=TRUE, message=FALSE}
cv_model <- ml_fit(cv, okc_train)
```

---

# Interoperability
Pipelines are fully interoperable with the other Spark APIs such as Python and Scala. 

## Save Pipeline Model

```{r, eval = FALSE, echo=TRUE, message=FALSE}
model_dir <- file.path("spark_model")
ml_save(cv_model$best_model, model_dir, overwrite = TRUE)
```

## Reconstruct the Model

```{r, eval = FALSE, echo=TRUE, message=FALSE}
model_reload <- ml_load(sc, model_dir)
```

## Retrieve a Stage from the Model

```{r, eval = FALSE, echo=TRUE, message=FALSE}
ml_stage(model_reload, "logistic_regression")
```


---

# Deployment
The process of taking a model and turning it into a service that others can consume.

## Batch
Processing many records at the same time. Execution time is not important as long it is reasonable.
`plumber`, `callr`, `httr`

## Real-Time
Scoring one or a few records at a time, but the latency is crucial (on the scale of <1 second).
`mleap`
---

# Questions 1 Preparation

1. Make a connection to the local spark cluster.
1. Load the `house_price.csv` file into Spark.
---

# Questions 2 EDA
1. Take a look at the data.
1. Check missing value for all features.
1. Drop the first column of the data.
1. Split data into train and test (size = 0.05) set.
1. Get numerical summaries of age and distance of the training data.
---

# Questions 3 Supervised Learning
1. Fit training data with decision tree.
1. Make prediction on test data.
1. Evaluate the prediction (report r2).
---

# Questions 4 Pipelines
1. Build a transformer for turning all features into vectors.
1. Build an estimator that estimates feature mean and standard deviation.
1. Create a pipeline with transformer, estimator, and decision tree.
1. Fit the pipeline model with training data.
1. Make prediction on test data using the pipeline model and evaluate the prediction (report r2).
1. Save pipeline model.
1. Reload pipeline model.
1. Extract one estimator from the pipeline.
1. Disconnect to the spark cluster
---

# Questions 5 Unsupervised Learning
1. Make a connection to the local spark cluster.
1. Load `iris` dataset into spark.
1. Take a look at the data.
1. Check missing value for chosen features (petal length, petal width).
1. Fit a model using k-means model with 3 clusters.
1. Predict the associated class.
1. Plot cluster membership.
1. Disconnect to the spark cluster.