---
title: "Advanced R"
subtitle: "Chapters 23-24"
author: "Ted Kyi"
institute: "OCRUG"
date: "2020-12-02"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, intro.css]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      
---
<style>
pre {
    display: block;
    font-family: monospace;
    white-space: pre;
    margin: 1em 0px;
    margin-top: 0em;
    margin-right: 0px;
    margin-bottom: 0em;
    margin-left: 0px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(profvis)
library(bench)
library(tibble)
```

# Measuring Performance

* R uses a sampling profiler
* Recommend visualizing using RStudio Profile menu, `profvis::profvis()`, or similar instead of `utils::Rprof()`
* Limitations/warnings
  * Can't profile C code
  * Anonymous functions are hard to track (you can temporarily name them)
  * Lazy evaluation can be confusing (you can temporarily use `force()`)

---
# Measuring Performance

* Microbenchmarking
* `bench::mark()` measures/compares time to run small snippets of code
* Focus on the minimum and median times (slow times can be caused by external things also running)

```{r bench example, echo=TRUE}
x <- runif(100)
(lb <- bench::mark(
  sqrt(x),
  x ^ 0.5
))
```

  
---
# Improving Performance

* Recommendations start with structuring your code to make optimization easy
* Search online for existing solutions
* Make your code do less -- skip nonessential parts
* Vectorize your code
* Avoid lots of copying of data


---
# Questions

### Chapter 23

1. 23.2.4 Question 1
2. 23.3.3 Question 2

### Chapter 24

1. 24.3.1 Question 2
2. 24.3.1 Question 3
3. 24.4.3 Question 1
4. 24.5.1 Question 1
5. 24.5.1 Question 3
