---
title: "Advanced R"
subtitle: "Chapter 25"
author: "John Peach"
institute: "OCRUG"
date: "2020-12-02"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, intro.css]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      
---
<style>
pre {
    display: block;
    font-family: monospace;
    white-space: pre;
    margin: 1em 0px;
    margin-top: 0em;
    margin-right: 0px;
    margin-bottom: 0em;
    margin-left: 0px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(lobstr)
library(tibble)
library(rlang)
library(purrr)
```

# When to use RCpp

* Loops that cannot be easily vectorised because subsequent iterations depend on previous ones.

* Recursive functions, or problems which involve calling functions millions of times. The overhead of calling a function in C++ is much lower than in R.

* Problems that require advanced data structures and algorithms that R does not provide. Through the standard template library (STL), C++ has efficient implementations of many important data structures, from ordered maps to double-ended queues.

---

# Writing Cpp Inline

```{r, echo=TRUE, eval=TRUE}
library(Rcpp)
cppFunction('int add(int x, int y, int z) {
  int sum = x + y + z;
  return sum;
}')
# add works like a regular R function
add
```

```{r, echo=TRUE}
add(1, 2, 3)
```

---

# R Type Vectors

The classes for the most common types of R vectors are: 
* `NumericVector`
* `IntegerVector`
* `CharacterVector`
* `LogicalVector`

* Scalars types
* `int`
* `double`
* `bool`
* `String`

Notice `char` is not supported

---

# Using sourceCpp

* use a stand alone `.cpp` file and include
```{cpp, echo=TRUE, eval=FALSE}
#include <Rcpp.h>
using namespace Rcpp;
```

For each function you want to export add

```{cpp, echo=TRUE, eval=FALSE}
// [[Rcpp::export]]
```

---

# Lists and Dataframes

* `List` and `Dataframes` are supported
* templates are supported for data type conversion

```{cpp, echo=TRUE, eval=FALSE}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
double mpe(List mod) {
  if (!mod.inherits("lm")) stop("Input must be a linear model");

  NumericVector resid = as<NumericVector>(mod["residuals"]);
  NumericVector fitted = as<NumericVector>(mod["fitted.values"]);

  int n = resid.size();
  double err = 0;
  for(int i = 0; i < n; ++i) {
    err += resid[i] / (fitted[i] + resid[i]);
  }
  return err / n;
}
```

* Also note the `.inherits()` method for assessing class information

---

# Functions

* The `Function` class supports functions

```{cpp, echo=TRUE, eval=FALSE}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
RObject callWithOne(Function f) {
  return f(1);
}
```

---

# Attributes

The following methods are used to access R attributes
* `.attrib()`
* `.names()`
* `.slot()` for S4 objects

```{r, echo=TRUE, eval=TRUE}
library(Rcpp)
cppFunction('NumericVector attribs() {
  NumericVector out = NumericVector::create(1, 2, 3);

  out.names() = CharacterVector::create("a", "b", "c");
  out.attr("my-attr") = "my-value";
  out.attr("class") = "my-class";

  return out;
}')
attribs()
```


* The `::create()` method allows a vector to be created

---

# Missing Values

Scaler values have special cases
* `NA_INTEGER` smallest integer, but still an integer
* `NA_STRING` Strings knows how to deal with missing values
* `NA_LOGICAL` NA are converted to TRUE in C++
* `NA_REAL` All logical expression of NaN return FALSE

Vectors support NA
```{r, echo=TRUE, eval=TRUE}
cppFunction('List missing_sampler() {
  return List::create(
    NumericVector::create(NA_REAL), IntegerVector::create(NA_INTEGER),
    LogicalVector::create(NA_LOGICAL), CharacterVector::create(NA_STRING));
}')
str(missing_sampler())
```

---

# STL - Interators

```{r, echo=TRUE, eval=FALSE}
double sum3(NumericVector x) {
  double total = 0;
  
  NumericVector::iterator it;
  for(it = x.begin(); it != x.end(); ++it) {
    total += *it;
  }
  return total;
}
```

* Create an iterator with `::iterator`
* First element is `.begin()`
* Last element is `.end()`
* increment with the standard `++`

---

# STL - Data Structures / Vectors

Has a huge number of data structures that will make your life easier

* `array`, `bitset`, `list`, `forward_list`, `map`, `multimap`, `multiset`, `priority_queue`, `queue`, `deque`, `set`, `stack`, `unordered_map`, `unordered_set`, `unordered_multimap`, `unordered_multiset`, and `vector`

`vector` class is templated and allows for efficient growth.

* `vector<int>`, `vector<bool>`, `vector<double>`, `vector<String>`
* Add new elements with `.push_back()`
* Preallocate space with `.reserve()`

---

# STL - Sets and Maps

Sets are collections of unique values
* Ordered `std::set<TYPE>` and unordered `std::unordered_set<TYPE>`
* It stores the presence or absence

Map like a set but stores additional date
* Ordered `std::map<KEY TYPE, VALUE TYPE>` and unordered `std::unordered_map<KEY TYPE, VALUE TYPE>`

---

# Questions

1. 25.2.6 Question 2

