---
title: "Advanced R"
subtitle: "Chapters 17-19"
author: "John Peach"
institute: "OCRUG"
date: "2020-11-11"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, intro.css]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      
---
<style>
pre {
    display: block;
    font-family: monospace;
    white-space: pre;
    margin: 1em 0px;
    margin-top: 0em;
    margin-right: 0px;
    margin-bottom: 0em;
    margin-left: 0px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(lobstr)
library(tibble)
library(rlang)
library(purrr)
```

# Code and Data

* `expr()` returns an expression. Used interactively
* `expr()` and `enquo()` are defusing functions (a.k.a. quoting)
* expressions are either a call, symbol, constant or pairlist
* `enexpr()` is used to capture user input

---

# Code is a tree

```{r, echo=TRUE}
lobstr::ast(f(a, "b"))
```

---

# Code can generate code

* Create on AST with `call2()`
```{r, echo=TRUE}
rlang::call2("+", 1, call2("*", 2, 3))
```

* Use `!!` to insert an expression into an AST
```{r, echo=TRUE}
xx <- expr(x + x)
yy <- expr(y + y)
expr(!!xx / !!yy)
```

---

# Evaluation run code

* `base::eval()` used to evaluate expressions
```{r, echo=yy}
x <- 10
y <- 100
eval(expr(x + y))
```

```{r, echo=TRUE}
string_math <- function(x) {
  e <- env(
    caller_env(),
    `+` = function(x, y) paste0(x, y),
    `*` = function(x, y) strrep(x, y)
  )
  eval(enexpr(x), e)
}
string_math("Hello " + "John")
string_math(("x" * 2 + "-y") * 3)
```

---

# Data mask

* `eval_tidy()` powers the tidy eval framework
```{r, echo=TRUE}
df <- data.frame(x=1:5, y = 11:15)
eval_tidy(expr(2 * x + y), df)
```

---

# Quosures

```{r, echo=TRUE}
with2 <- function(df, expr) {
  a <- 1000
  eval_tidy(enexpr(expr), df)
}
df <- data.frame(x = 1:3)
a <- 10
with2(df, x + a)
```

```{r, echo=TRUE}
with2 <- function(df, expr) {
  a <- 1000
  eval_tidy(enquo(expr), df)
}
with2(df, x + a)
```

* Always use `enquo()` with data masks
* `expr()` defuses your own expression
* `enquo()` defuses expressions supplied as argument by the user of a function

---

# Symbols

* A symbol represents the name of an object
* `rlang::sym()` turn a string into a symbol
```{r, echo=TRUE}
expr(x)
sym("x")
```

Convert a symbol back into a string
```{r, echo=TRUE}
as_string(sym("x"))
```

---

# Calls

* A call object represents a captured function call
* Calls behave like lists

```{r, echo=TRUE}
x <- expr(read.table("important.csv", row.names = FALSE))
ast(!!x)
```

Standardize the arguments
```{r, echo=TRUE}
rlang::call_standardise(x)
ast(!!rlang::call_standardise(x))
```

---

# Parsing

* `rlang::parse_expr()` Convert text into code
```{r, echo=TRUE}
x <- rlang::parse_expr("y <- x + 10")
is_call(x)
x
```

* `rlang::parse_exprs()` returns a list of expressions
* deparse with `rlang::expr_text()`
```{r, echo=TRUE}
expr_text(x)
```

---

# Quasiquotation

* quotation is the act of capturing an unevaluated expression
* unquotation is the ability to selectively evaluate parts of an quoted expression
* quasiquotation is quotation and unquotation
* In tidy evaluation, all quoting functions are actually quasiquoting functions because they also support unquoting
* tidy evaluation consists of:
  1. quasiquotation
  1. quasures
  1. data mask
* The distinction between quoted and evaluated arguments is important:
  * An evaluated argument obeys R’s usual evaluation rules.
  * A quoted argument is captured by the function, and is processed in some custom way.
  
---

# Unquoting

* `!!` unqoute an expression

Unquoting a function, wrap it in ( )
```{r, echo=TRUE}
f <- expr(pkg::foo)
expr((!!f)(x, y))
```

Unquoting a missing argument
```{r, echo=TRUE, eval=FALSE}
arg <- missing_arg()
expr(foo(!!arg, !!arg))
```
Above is an error, use `rlang::maybe_missing()`
```{r, echo=TRUE}
arg <- missing_arg()
expr(foo(!!maybe_missing(arg), !!maybe_missing(arg)))
```

---
# Unquoting

Unquoting special forms
* `expr(df$!!x)` creates an error
```{r, echo=TRUE}
x <- expr(x)
expr(`$`(df, !!x))
```

---

# Unquote-splice

* `!!!` unquote many arguments
```{r, echo=TRUE}
xs <- exprs(1, a, -b)
exprs(f(!!!xs, y))
```

Use `!!!` when the parameters are already in a list (a.k.a. spatting)
```{r, echo=TRUE}
dfs <- list(
  a = data.frame(x = 1, y = 2),
  b = data.frame(x = 3, y = 4)
)
dplyr::bind_rows(!!!dfs)
```

---

# Unquoting on the left-hand side

Use `:=` when unquoting on the LFS of an evaluation
```{r, echo=TRUE}
var <- "x"
val <- c(4, 3, 9)
tibble::tibble(!!var := val)
```

---

# Calling a function that does not support tidy dots

Use `exec`. This looks to work just like `do.call()`
```{r, echo=TRUE}
arg_name <- "na.rm"
arg_val <- TRUE
exec("mean", 1:10, !!arg_name := arg_val)
```
Note the use of `!!` and `:=`

Or use a list with `!!!`
```{r, echo=TRUE}
args <- list(x = 1:10, na.rm = TRUE, trim = 0.1)
exec("mean", !!!args)
```

---

# `dots_list`

* Used to control lists. It allows for empty elements and homonyms
* `list2()` is a wrapper around `dots_list()` with common defaults.
* `dots_list()` has more control over behaviour of elements in the list

```{r, echo=TRUE}
dots_list(
  y = 1:5,
  z = 3:-1,
  x = 5:1,
)
```


Note the comma on the last line. Normally this would be an error.

---

# Questions

1. 18.2.4 Exercise Question 1
1. 18.3.5 Exercise Question 1
1. 18.3.5 Exercise Question 4
1. 18.3.5 Exercise Question 5
1. 18.4.4 Exercise Question 1
1. 18.4.4 Exercise Question 9
1. 19.2.2 Exercise Question 2
1. 19.3.6 Exercise Question 2
1. 19.3.6 Exercise Question 4
1. 19.4.8 Exercise Question 1
1. 19.6.5 Exercise Question 3
