---
title: "Chap_4-5_exercise_template"
author: "Your Name"
date: "2/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Step 1. Preparation

## 1.1 Load libraries
```{r}
library(sparklyr)
library(ggplot2)
library(dbplot)
library(dplyr)
```



What are the three major trends in [cluster computing](https://therinspark.com/clusters.html#clusters-overview)?
```{r, eval = FALSE, echo=TRUE, message=FALSE}

```

What are (some) attributes of [on-premise clusters](https://therinspark.com/clusters.html#on-premise)?
```{r, eval = FALSE, echo=TRUE, message=FALSE}

```

What are (some) benefits of [cloud computing clusters](https://therinspark.com/clusters.html#cloud)?
```{r, eval = FALSE, echo=TRUE, message=FALSE}

```

Can someone [explain Kubernetes](https://therinspark.com/clusters.html#kubernetes) to me like I'm a five year old?
```{r, eval = FALSE, echo=TRUE, message=FALSE}

```

What is the purpose of [cluster managers](https://therinspark.com/clusters.html#clusters-manager) in Spark?
```{r, eval = FALSE, echo=TRUE, message=FALSE}

```

What [cluster managers](https://therinspark.com/clusters.html#clusters-manager) are available in Spark?
```{r, eval = FALSE, echo=TRUE, message=FALSE}

```

What [distributions](https://therinspark.com/clusters.html#distributions) are available for Spark?
```{r, eval = FALSE, echo=TRUE, message=FALSE}

```

What are some [common tools](https://therinspark.com/clusters.html#tools) available for Spark?
```{r, eval = FALSE, echo=TRUE, message=FALSE}

```


## Start a [cluster manager](https://therinspark.com/clusters.html#clusters-manager) of your choice - AWS?
```{r, eval = FALSE, echo=TRUE, message=FALSE}


```

## Retrieve the Spark installation directory
```{r, eval = FALSE, echo=TRUE, message=FALSE}


```


## How do you find installed versions of Spark?
```{r, eval = FALSE, echo=TRUE, message=FALSE}
 


```

## How do you specify [SPARK_HOME](https://therinspark.com/connections.html#connections-spark-home) environment variable in your cluster
```{r, eval = FALSE, echo=TRUE, message=FALSE}


```

## How do you make a [standalone connection](https://therinspark.com/connections.html#connections-standalone?
```{r, eval = FALSE, echo=TRUE, message=FALSE}


```

## Start a worker node
```{r, eval = FALSE, echo=TRUE, message=FALSE}

```




## 5.8 Disconnect to the spark cluster.
```{r}


```

