---
title: "chapter 21 - 24"
author: "Xi Chen"
date: "2/19/2020"
output: ioslides_presentation
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
```{r include = FALSE,echo = FALSE}
library(tidyverse)
library(ggplot2)
```

## Chapter 21 
#### R Markdown
* integrate prose, code and results together
* output to multiple formats, pdf, word, html, slides
* for easy communication and collaboration 

#### Text formatting 
* italic: *italic*  
* bold: **bold**  
* inline code(backtick): `code`
* headings (hash #): 
* lists(bullet lists, numbered list):
* link and images(pointy brackets, path to the file): 

## Chapter 21 R markdown
#### Code chunk
* Cmd/Ctrl + Alt + I
* special chunk name: setup

#### chunk option
* eval = FALSE    ----> example,disable a large chunk
* include = FALSE ----> setup, hide code, hide result
* echo = FALSE    ----> report, hide code, show result
* message = FALSE or warning = FALSE 
* results = 'hide' 
* fig.show = 'hide' hides plots.
* error = TRUE 

## Chapter 21
#### table
* knitr::kable

#### caching
* cache = TRUE
* dependson = "dataset name"

#### global options
* knitr::opts_chunk$set(
        echo = FALSE
)

#### parameters
* params: 
    my_class: "suv"
* class == params$my_class

## Chapter 22 exploratory  -> expository 
##### labels
* labs(  
 x = "Engine displacement (L)",  
 y = "Highway fuel economy (mpg)",  
 colour = "Car type")
* quote() to use math equations  
`quote(sum(x[i] ^ 2, i == 1, n))`    
`quote(alpha + beta + frac(delta, theta))`

## Chapter 22
##### annotations
* geom_text(aes(label = model), data = best_in_class)
* geom_label(aes(label = model), data = best_in_class, nudge_y = 2, alpha = 0.2)
* ggrepel::geom_label_repel(aes(label = model), data = best_in_class)
* stringr::str_wrap(width = 40)
* geom_hline(), geom_vline()
* geom_rect(), geom_segment with arrow argument

## Chapter 22
#### scales
* scale_x_continuous(breaks = seq(15,40,by=5))
* scale_x_date(breaks=presidential$start, date_labels="%y")
* theme(legend.position = "bottom") +
  guides(colour=guide_legend(nrow=1, override.aes = list(size=4)))
* scale_x_log10() + scale_y_log10()
* scale_colour_brewer(palette = "Set1")
* scale_colour_manual(values = c(Republican = "red", Democratic = "blue"))

## Chapter 22
#### zooming
```{r echo = TRUE, fig.show='hide', message=FALSE}
ggplot(mpg, mapping = aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth() +
  coord_cartesian(xlim = c(5, 7), ylim = c(10, 30))

mpg %>%
  filter(displ >= 5, displ <= 7, hwy >= 10, hwy <= 30) %>%
  ggplot(aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth()
```
## Chapter 22
#### zooming
```{r echo = TRUE, fig.show='hide', message=FALSE}
suv <- mpg %>% filter(class == "suv")
compact <- mpg %>% filter(class == "compact")
x_scale <- scale_x_continuous(limits = range(mpg$displ))
y_scale <- scale_y_continuous(limits = range(mpg$hwy))
col_scale <- scale_colour_discrete(limits = unique(mpg$drv))

ggplot(suv, aes(displ, hwy, colour = drv)) +
  geom_point() +
  x_scale +
  y_scale +
  col_scale

ggplot(compact, aes(displ, hwy, colour = drv)) +
  geom_point() +
  x_scale +
  y_scale +
  col_scale
```
## Chapter 22
#### themes
* theme_classic()

#### save your plots
* ggsave("my-plot.pdf")

#### figure sizing
* fig.width,fig.height, fig.asp
* out.width, out.height
* fig.show = "hold"
* fig.cap inline -> floating?
* dev="png"

## Chapter 23
#### output
* render to multiple outputs by supllying a list of formats
* code_folding:hide (code hidden by default, visible by click)
* new slide begins with first level(#), second(##), or horizontal rule(***)

#### dashboard
* ('#') begins a new page
* (##) begins a new column
* (###) begins a new row

#### interactivity
* htmlwidgets
* shiny, websites
* bookdown 

## Good practice
* add a descriptive title
* date your documents
* note the failure as well
* write code to correct the data
* knit the notebook before finish the day
* track the versions of the packages
* store them in individual projects

## Questions
* how to put figures side by side
* start new line without using list
* exercise 28.3.3
* exercise 28.4.2
* exercise 27.4.4





